---
title: "HW5"
author: "Xinyue Lu"
date: "2020/2/25"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```



#### 7.9 Exercise 6
In this exercise, you will further analyze the Wage data set considered throughout this chapter.
```{r}
# Load Data
library(ISLR)
attach(Wage)
```
##### (a) Perform polynomial regression to predict wage using age. Use cross-validation to select the optimal degree d for the polynomial. What degree was chosen, and how does this compare to the results of hypothesis testing using ANOVA? Make a plot of the resulting polynomial ﬁt to the data. 

The plot of cross validation error and degree showed that the error dropped sharply at degree of 2 and changed little after the degree of 4.Model with 2 polynomial degree or 3 polynomial degree should be choosed. The ANOVA model choosed the model with 3 polynomial degree. 
```{r}
library(boot)
attach(Wage)
# Choose by Cross Validation
error = rep(0, 10)
for(i in 1:10){
fit=glm(wage~poly(age,i),data=Wage)
set.seed(666)
error[i]=cv.glm(Wage,fit,K=10)$delta[2]
}
plot(x=seq(1,10), error, xlab="Degree", ylab="Cross Validation Error", type="l")
# Choose by ANOVA
fit.1=lm(wage~age,data=Wage)
fit.2=lm(wage~poly(age,2),data=Wage)
fit.3=lm(wage~poly(age,3),data=Wage)
fit.4=lm(wage~poly(age,4),data=Wage)
fit.5=lm(wage~poly(age,5),data=Wage)
anova(fit.1,fit.2,fit.3,fit.4,fit.5)
coef(summary(fit.5))

```

##### (b) Fit a step function to predict wage using age, and perform crossvalidation to choose the optimal number of cuts. Make a plot of the ﬁt obtained.

The optimal number of cuts was 8.
```{r}
error = rep(0, 10)
for(i in 2:10){
Wage$age.cut=cut(Wage$age,i)  
fit=glm(wage~age.cut,data=Wage)
set.seed(666)
error[i]=cv.glm(Wage,fit,K=10)$delta[2]
}
plot(x=seq(2,10), error[-1], xlab="Degree", ylab="Cross Validation Error", type="l")

fit = glm(wage~cut(age, 8), data=Wage)
agelims = range(Wage$age)
age.grid = seq(from=agelims[1], to=agelims[2])
pred = predict(fit, data.frame(age=age.grid))
plot(wage~age, data=Wage)
lines(age.grid, pred, col="red")
```



#### 1. Run the lab 7.8 pages 287-291, just for Wage as a response 

#### 1.1 Polynomial Regression and Step Functions
#### 1.1.1 Fit the model with different expressions
```{r}
fit=lm(wage~poly(age,4),data=Wage)
coef(summary(fit))
# 3 methods produce the same coefficient estimates
fit2=lm(wage~poly(age,4,raw=T),data=Wage)
coef(summary(fit2))
fit2a=lm(wage~age+I(age^2)+I(age^3)+I(age^4),data=Wage)
coef(fit2a)
fit2b=lm(wage~cbind(age,age^2,age^3,age^4),data=Wage)
coef(fit2b)
```
#### 1.1.2 Plot the predicted values
```{r}
agelims=range(age)
age.grid=seq(from=agelims[1],to=agelims[2])
par(mfrow=c(1,2),mar=c(4.5,4.5,1,1),oma=c(0,0,4,0))
# Plot fit
preds=predict(fit,newdata=list(age=age.grid),se=TRUE)
se.bands=cbind(preds$fit+2*preds$se.fit,preds$fit-2*preds$se.fit)
plot(age,wage,xlim=agelims,cex=.5,col="darkgrey")
title("Degree-4 Polynomial (Right:Raw)",outer=T)
lines(age.grid,preds$fit,lwd=2,col="blue")
matlines(age.grid,se.bands,lwd=1,col="blue",lty=3)
# Plot fit2
preds2=predict(fit2,newdata=list(age=age.grid),se=TRUE)
se.bands2=cbind(preds2$fit+2*preds2$se.fit,preds2$fit-2*preds2$se.fit)
plot(age,wage,xlim=agelims,cex=.5,col="darkgrey")
lines(age.grid,preds2$fit,lwd=2,col="blue")
matlines(age.grid,se.bands2,lwd=1,col="blue",lty=3)
# Maximum difference between prediction of fit and fit2
max(abs(preds$fit-preds2$fit))
```

#### 1.1.3 Use ANOVA to determine the degree of the model

The p-value of the model fit.4 was greater than 0.5, which meant that model fit.4 and model fit.3 were similar in performance. But model fit.2 was not so good as model fit.3 because the p-value of model fit.3 was smaller than 0.5.
```{r}
fit.1=lm(wage~age,data=Wage)
fit.2=lm(wage~poly(age,2),data=Wage)
fit.3=lm(wage~poly(age,3),data=Wage)
fit.4=lm(wage~poly(age,4),data=Wage)
fit.5=lm(wage~poly(age,5),data=Wage)
anova(fit.1,fit.2,fit.3,fit.4,fit.5)
coef(summary(fit.5))
```
#### 1.1.4 ANOVA can be used to compare different models
```{r}
fit.1=lm(wage~education+age,data=Wage)
fit.2=lm(wage~education+poly(age,2),data=Wage)
fit.3=lm(wage~education+poly(age,3),data=Wage)
anova(fit.1,fit.2,fit.3)
```
#### 1.1.5 Fit a Polinomial Logistic Regression
```{r}
fit=glm(I(wage>250)~poly(age,4),data=Wage,family=binomial)
preds=predict(fit,newdata=list(age=age.grid),se=T)
pfit=exp(preds$fit)/(1+exp(preds$fit))
se.bands.logit = cbind(preds$fit+2*preds$se.fit, preds$fit-2*preds$se.fit)
se.bands = exp(se.bands.logit)/(1+exp(se.bands.logit))

plot(age,I(wage>250),xlim=agelims,type="n",ylim=c(0,.2))
points(jitter(age), I((wage>250)/5),cex=.5,pch="|",col="darkgrey")
lines(age.grid,pfit,lwd=2, col="blue")
matlines(age.grid,se.bands,lwd=1,col="blue",lty=3)
```

preds=predict(fit,newdata=list(age=age.grid),type="response",se=T)



```{r}
table(cut(age,4))
fit=lm(wage~cut(age,4),data=Wage)
coef(summary(fit))
```


#### 1.2 Fit the Model with Splines
#### 1.2.1 Cubic Splines
```{r}
library(splines)
fit=lm(wage~bs(age,knots=c(25,40,60)),data=Wage)
pred=predict(fit,newdata=list(age=age.grid),se=T)
plot(age,wage,col="gray")
lines(age.grid,pred$fit,lwd=2)
lines(age.grid,pred$fit+2*pred$se,lty="dashed")
lines(age.grid,pred$fit-2*pred$se,lty="dashed")
dim(bs(age,knots=c(25,40,60)))
dim(bs(age,df=6))
attr(bs(age,df=6),"knots")
```

#### 1.2.2 Natural Splines
```{r}
fit2=lm(wage~ns(age,df=4),data=Wage)
pred2=predict(fit2,newdata=list(age=age.grid),se=T)
plot(age,wage,col="gray")
lines(age.grid, pred2$fit,col="red",lwd=2)
```
#### 1.2.3 Smooth Splines
```{r}
plot(age,wage,xlim=agelims,cex=.5,col="darkgrey")
title("Smoothing Spline")
fit=smooth.spline(age,wage,df=16)
fit2=smooth.spline(age,wage,cv=TRUE)
fit2$df
lines(fit,col="red",lwd=2)
lines(fit2,col="blue",lwd=2)
legend("topright",legend=c("16 DF","6.8 DF"),col=c("red","blue"),lty=1,lwd=2,cex=.8)
```
#### 1.2.4 Local Regression
```{r}
plot(age,wage,xlim=agelims,cex=.5,col="darkgrey")
title("Local Regression")
fit=loess(wage~age,span=.2,data=Wage)
fit2=loess(wage~age,span=.5,data=Wage)
lines(age.grid,predict(fit,data.frame(age=age.grid)),col="red",lwd=2)
lines(age.grid,predict(fit2,data.frame(age=age.grid)),col="blue",lwd=2)
legend("topright",legend=c("Span=0.2","Span=0.5"),col=c("red","blue"),lty=1,lwd=2,cex=.8)
```



```{r}
x = seq(-2,2,length=100)
y = 1+x+(-2)*(x-1)^2*I(x>1)
plot(x, y,type='l')
```

```{r}
x = seq(-2,2,length=100)
y = 1+I(0<=x)*I(x<=2)-(x-1)*I(1<=x)*(x<=2)+3*((x-3)*I(3<=x)*I(x<=4)+I(4<=x)*I(x<=5))
plot(x, y,type='l')
```

