---
title: "HW6"
author: "Xinyue Lu"
date: "2020/3/3"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


##### Implement the Nadaraya-Watson kernel regression based on the Epanechnikov and Gaussian kernels. Use rule of thumb and cross-validation to determine the bandwidth parameter h. For cross validation,implement grid search around the rule of thumb estimate. Compare the results. The data set is Wage, you can find it in the library(ISLR) use attach(Wage)

Normal Kernel by ksmooth

First, try 3 bandwidths 1, 5 and 10. Tt showed that bandwidth 5 produce the most preferable estimates, because the other two were either too rough or too smooth. Next, use cross validation to search for the best bandwidths from 5 to 9 by 0.2 a step. The best bandwidth was 7, which produced the smallest MSE 1595.884.
```{r}
library(ISLR)
attach(Wage)

# set cross validation set
set.seed(666)
idx = sample(seq(1:3000))
index = cbind(idx[1:600],idx[601:1200],idx[1201:1800],idx[1801:2400],idx[2401:3000])

# choose the bandwidth by plots
test = index[,1]
train = (-test)
train.age = age[train] 
train.wage = wage[train] 
test.age = age[test] 
test.wage = wage[test] 
# bandwidth =1
pred = ksmooth(train.age, train.wage, kernel = "normal", bandwidth = 1, range.x = range(train.age),
        n.points = length(train.age),x.points = test.age)
plot(test.age,test.wage)
lines(pred$x,pred$y,col="red")
title("Normal Kernel with Bandwidth=1")
# bandwidth =5
pred = ksmooth(train.age, train.wage, kernel = "normal", bandwidth = 5, range.x = range(train.age),
        n.points = length(train.age),x.points = test.age)
plot(test.age,test.wage)
lines(pred$x,pred$y,col="red")
title("Normal Kernel with Bandwidth=5")
# bandwidth =10
pred = ksmooth(train.age, train.wage, kernel = "normal", bandwidth = 10, range.x = range(train.age),
        n.points = length(train.age),x.points = test.age)
plot(test.age,test.wage)
lines(pred$x,pred$y,col="red")
title("Normal Kernel with Bandwidth=10")
 
#############################################
# use cross validation to find best bandwidth
mse = matrix(0,20,5)
for(h in 1:20){
b=h*0.2+5
for(j in (1:5)){
test = index[,j]
train = (-test)
train.age = age[train] 
train.wage = wage[train] 
test.age = age[test] 
test.wage = wage[test] 

#fit = ksmooth(train.age, train.wage, kernel = "normal", bandwidth = b, range.x = range(train.age), n.points = length(train.age))
#plot(train.age,train.wage)
#lines(fit$x,fit$y,col="red")

pred = ksmooth(train.age, train.wage, kernel = "normal", bandwidth = b, range.x = range(train.age),
        n.points = length(train.age),x.points = test.age)
#plot(test.age,test.wage)
#lines(pred$x,pred$y,col="red")

mat = cbind(test.age,test.wage)
o = order(mat[,"test.age"])
mat = mat[o,]
mse[h,j]=mean((mat[,"test.wage"]-pred$y)^2)
}}

# MSE of the cross validation with bandwidth [5 to 9]
MSE = rowMeans(mse)
MSE
min(MSE)
# Find the Bandwidth with Minimal MSE
index = which(MSE == min(MSE))
best.bandwidth = index*0.2+5
best.bandwidth
pred = ksmooth(train.age, train.wage, kernel = "normal", bandwidth = 6.8, range.x = range(train.age),
        n.points = length(train.age),x.points = test.age)
plot(test.age,test.wage)
lines(pred$x,pred$y,col="red")
title("Normal Kernel with Bandwidth=6.8")
```


